---
title: "Day09: data structures"
author: "csiu"
date: "2015-10-06"
output:
    html_document:
        keep_md: yes
---

- `dplr` - domain specific language
- [lecture outline](http://stat545-ubc.github.io/cm009_r-objects-collections.html)

+-----------+---------------+-----------+-----------+
| "flavor"  | type reported | mode()    | class()   |
|           | by typeof()   |           |           |
+===========+===============+===========+===========+
| character | character     | character | character |
+-----------+---------------+-----------+-----------+
| logical   | logical       | logical   | logical   |
+-----------+---------------+-----------+-----------+
| numeric   | integer       | numeric   | integer   |
|           | or double     |           | or double |
+-----------+---------------+-----------+-----------+
| factor    | integer       | numeric   | factor    |
+-----------+---------------+-----------+-----------+

### intrinsic vector nature of R
```{r results="hide"}
x <- 3 * 4   ## [1] 12
is.vector(x) ## [1] TRUE
length(x)    ## [1] 1


x[2] <- 100  ## [1]  12 100
x[5] <- 3    ## [1]  12 100  NA  NA   3
x[11]        ## [1] NA
x[0]         ## numeric(0)
```

```{r}
x <- 1:4

#======#
# GOOD #
#======#
y <- x^2

#=====#
# BAD #
#=====#
## or the for loop version?
z <- vector(mode = mode(x), length = length(x)) #intialize: should be mode of x and length of x
for(i in seq_along(x)) {
  z[i] <- x[i]^2 #assign
}

#===============#
identical(y, z)
```

### mean and standard deviation of random normal variates can be provided as vectors
```{r results="hide"}
set.seed(1999)
rnorm(5, mean = 10^(1:5))
round(rnorm(5, sd = 10^(0:4)), 2)
```

### R also recycles vectors
```{r}
y <- 1:3 ## 1 2 3
z <- 3:7 ## 3 4 5 6 7
y + z
```
... but if one variable is a factor of another, then get no warnings

### Indexing a vector
Most common, useful ways to index a vector

- logical vector: keep elements associated with TRUEs, ditch the FALSEs
- vector of positive integers specifying the keepers
- vector of negative integers specifying the losers
- character vector, naming the keepers

### lists hold just about anything
- data.frames are lists!

### Indexing arrays, e.g. matrices
- array & matrices ... each element has to be the same flavor

- vector vs list
- matrix vs data.frame

```{r error=TRUE}
data.frame(x=1:3, y=1:6) ##recycling

dplyr::data_frame(x=1:3, y=1:6) ##dplyr does not allow recycling
dplyr::data_frame(x= rep(1:3, 2), y=1:6) ##need to explicitly do it
```

> The drop = argument controls whether the return value should be an atomic vector (`drop = TRUE`) or a matrix with a single row or column (`drop = FALSE`). 

```{r}
jMat <- outer(as.character(1:4), as.character(1:4),
              function(x, y) {
                paste0('x', x, y)
                })
jMat[ , 3]
jMat[ , 3, drop = FALSE] # getting column 3
```
