---
title: "functions"
author: "csiu"
date: "October 8, 2015"
output: 
    html_document:
        keep_md: yes
---
```{r}
library(gapminder)
```

## How to write functions in R

- [write your own function](http://stat545-ubc.github.io/block011_write-your-own-function-01.html)
- for use in split-apply-combine

### Example: Range between max and min
```{r results='hide'}
## Before
max(gapminder$lifeExp) - min(gapminder$lifeExp)
diff(range(gapminder$lifeExp))

## After
max_minus_min <- function(x){
  return(max(x) - min(x))
}

max_minus_min(gapminder$lifeExp)
```

- Should check function to see if it works
```{r}
max_minus_min(1:10)

max_minus_min(runif(100)) #random number from 0 to 1 .. should be close to 1
```

- Better to make errors more readable
    - use assertion/runtime assertion
    - checked against expectation/claim

```{r error=TRUE}
max_minus_min(gapminder$continent)

max_minus_min <- function(x){
  if (!is.numeric(x)) {
    stop("Sorry, but function only works for numeric input.\n",
         "You have provided an object of class: ",
         class(x))
  }
  max(x) - min(x)
}

max_minus_min(gapminder$continent)

## or, to be more concise ...
mmm <- function(x){
  stopifnot(is.numeric(x))  
  max(x) - min(x)
}
```

### Resources
> Packages for runtime assertions (the last 3 seem to be under more active development than assertthat):
> 
> - `assertthat` on CRAN and GitHub the Hadleyverse option
> - `ensurer` on CRAN and GitHub general purpose, pipe-friendly
> - `assertr` on CRAN and GitHub explicitly data pipeline oriented
> - `assertive` on CRAN and Bitbucket rich set of built-in functions
> 
> Hadley Wickham’s book Advanced R


### Example: quantiles

```{r}
quantile(gapminder$lifeExp)
the_quantiles <- quantile(gapminder$lifeExp, probs=c(0.25, 0.75))
max(the_quantiles) - min(the_quantiles)
IQR(gapminder$lifeExp)
```

The function:
```{r}
qdiff1 <- function(x, probs) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x, probs=probs)
  max(the_quantiles) - min(the_quantiles)
}
qdiff1(gapminder$lifeExp, c(0.25, 0.75))
```
- use arguments which make sense
