---
title: "stat547"
author: "csiu"
date: "October 27, 2015"
output: html_document
---
8,13,9,9,11 - 10:15

- [syllabus](http://stat545-ubc.github.io/syllabus.html)
- [lecture outline](http://stat545-ubc.github.io/cm101_files-out-in.html)

- David Ng -- [Candy Survey](https://boingboing.net/2015/10/23/only-you-can-determine-what-to.html) from Boing Boing

## Today: Reading data in/out of file
- Today's output is tomorrow's input
- Don't be too cute/clever (e.g. talk to excel in 3 years from now)

```{r error=TRUE}
## FIND LOCATION OF FILE IN PACKAGE
(gap_tsv <- system.file("gapminder.tsv", package="gapminder"))

## LOAD FILE
gapminder <- read.table(gap_tsv, header=TRUE, sep="\t", quote="")

assertthat::assert_that(nrow(gapminder) == 1700)
```

- [`readr::`](https://github.com/hadley/readr)
    - default mentality of not turning string as factor
    - be tbl_df by default
    - argument namees more consistent
    - noticeable faster (for intermediate file size realm)
    - encoding issue works out nicely

```{r}
library(readr)
gapminder <- read_tsv(gap_tsv)
str(gapminder)
```

- `saveRDS()` and `readRDS()` - Serialization Interface for Single Objects
    - factor levels order is retained
    
- `dput()` and `dget()` - Write an Object to a File or Recreate it
    - plain text representation 
    
    > Writes an ASCII text representation of an R object to a file or connection, or uses one to recreate the object.

## Free play

- [Candy survey](https://github.com/STAT545-UBC/Discussion/issues/193)
    - RE: https://github.com/STAT545-UBC/Discussion/issues/193
- clean up gapminder data
      
      
    
