---
title: "dplyr"
author: "csiu"
date: "2015-09-29"
output:
    html_document:
        keep_md: yes
---

- [lecture outline](http://stat545-ubc.github.io/cm007_dplyr-intro.html)
- theme of the week: `dplyr`
- split-apply-combine workflow

## dplyr
- [Intro to dplyr](http://stat545-ubc.github.io/block009_dplyr-intro.html): data.frame in, data.frame out
    - good for subsetting
- older package = `plyr`

```{r}
suppressPackageStartupMessages(library(dplyr)) ## dplyr is chatty
library(gapminder)
library(ggplot2)

## what packages is attached
search()
```

```{r}
## transform to "tbl_df"" object
(gtbl <- tbl_df(gapminder))

## get a glimpse of data
glimpse(gtbl)

## special: 
class(gtbl)
```
- `tbl_df` -special class of-> `tbl` -special class of-> `data.frame`

### `filter()`: row-wise filtering (and `select()`: column-wise filtering)
```{r}
## [WHAT YOU WOULD HAVE DONE BEFORE]
snippet <- subset(gapminder, country == "Canada")
#     country continent year lifeExp      pop gdpPercap
# 241  Canada  Americas 1952  68.750 14785584  11367.16
# 242  Canada  Americas 1957  69.960 17010154  12489.95
# 243  Canada  Americas 1962  71.300 18985849  13462.49
# ...

## [WHAT YOU WOULD DO AFTER]
filter(gtbl, country %in% c("Canada"), lifeExp < 70)
```
- with `dplyr::filter` (and other functions), don't need "dataframe`$`variable"
    - i.e. assumes look within data.frame
- multiple expression in function get logical-and

### pipe operator: `%>%`
- new, mildly controversial, from package `magrittr`
- Shortcut: `command-shift-m`

```{r fig.path='figure/cm007-'}
## Example
gapminder %>% 
  filter(continent == "Americas") %>% 
  select(country, lifeExp) %>%
  
  #head(50) %>% tail(10) %>% 
  
  ggplot(aes(x=country, y=lifeExp)) + ## can pipe into ggplot
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### List of functions from `dplyr`
- `filter()`: Return **rows** with matching conditions
- `select()`: Select/rename **variables** by name

- `mutate()`: Add new variables
- `arrange()`: Arrange rows by variables
- `rename()`: Select/rename variables by name

- `group_by()`: Group a tbl by one or more variables