---
title: "Day08: dplyr part2 + R data structures"
author: "csiu"
date: "2015-10-01"
output:
    html_document:
        keep_md: yes
---
```{r include=FALSE}
knitr::opts_chunk$set(fig.path='figure/cm008-')
```

- [lecture outline](http://stat545-ubc.github.io/cm008_dplyr-contd.html)

- R package for captioning: [captioner](https://cran.r-project.org/web/packages/captioner/index.html)
- use rawgit.com to view HTML

## dplyr part2
- [dplyr functions for a single dataset](http://stat545-ubc.github.io/block010_dplyr-end-single-table.html)

```{r}
suppressPackageStartupMessages(library(dplyr))
library(gapminder)
gtbl <- gapminder %>%
  tbl_df
```

### `mutate()`: to add new variable
- like `transform` in base R

### `arrange()`: to arrange row order
```{r}
gtbl %>%
  arrange(year, country) ## order by year and then country
```

### `rename()`: to rename variables 
- i.e. `lifeExp` to `life_exp`
```{r}
gtbl %>%
  rename(life_exp = lifeExp,
         gdp_percap = gdpPercap) %>% 
  head(4)
```

### `group_by()`: Group a tbl by one or more variables
- split-apply-combine
- group_by will group rows

- `summarize()`/`summarise()`: n observations -> distill to 1 observation
- `window()`: n observations -> maps to n observations
- `do()` (will cover later)

#### Example: Counting things inside each group with `summarize()`
```{r}
gtbl %>% 
  group_by(continent) %>% 
  summarize(n_obs = n())

table(gtbl$continent)
```
- _Point: Don't want super important information to be names_

#### `summarize_each()`: Summarise and mutate multiple columns
> ... applies the same summary function(s) to multiple variables. Let’s compute average and median life expectancy and GDP per capita by continent by year … but only for 1952 and 2007.
```{r}
gtbl %>%
  filter(year %in% c(1952, 2007)) %>%
  group_by(continent, year) %>%
  summarise_each(funs(mean, median), lifeExp, gdpPercap) %>% 
  head(4)
```

#### Window functions
- `min_rank()` - ranks list according to ascending order; ties are broken my taking the min
- `desc()` - transforms vector to format that will be sorte in descending order
```{r}
gtbl %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  arrange(year) %>%
  group_by(year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>%  #find min or max
  head(4)
```
- alternative approach: `top_n()` which wraps `filter` and `min_rank` to select top n entries in each group
```{r}
## ANOTHER EXAMPLE
gtbl %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  arrange(year) %>%
  group_by(year) %>%
  #top_n(1)               ## gets the min
  top_n(1, desc(lifeExp)) %>% ## gets the max
  head(4)
```

----
## Jenny's example code:
```{r}
library(ggplot2)
library(gapminder)

ggplot(gapminder, aes(x = continent, y = gdpPercap)) +
  scale_y_log10() +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4) +
  stat_summary(fun.y = min, colour = "blue", geom = "point", size = 5) +
  stat_summary(fun.y = max, colour = "orange", geom = "point", size = 5)
```