---
title: "Build R package"
author: "csiu"
date: "November 10, 2015"
output: 
    html_document:
        keep_md: yes
---

- [lecture outline](http://stat545-ubc.github.io/cm105_packages.html)

- resource for R packages by Hadley Wickham: http://r-pkgs.had.co.nz

- "install package into library"

```{r}
R.home() ## main directory where R lib is

.Library      ## default library which ships with R
dir(.Library) ## the packages which ships with R

.libPaths() ## other library path; default is first one

# readLines("~/.Renviron")
# R_LIBS=       ## path of other libs
# GITHUB_TOKEN= ## access github API from R
# GITHUB_PAT=   ## ditto
# NOT_CRAN=true
```

- Content of R package
    - `man/` 
    - `R/` - .R files
    - `src/` - c/c++/other scripts
    - `test/`
    - `inst` - catch all ... files which ship with package

- How do installed packages get into memory
    - Build & Reload
    - `load_all()`
    - `library()`, for when package live in default library
    
- Your first devtools

| function | info |
|:--------:|:----:|
| `devtools::create()` | setup a new package |
| `devtools::document()` | wrapper that uses `roxygen2` to make formal documentation and NAMESPACE | 
| RStudio Build & Reload ; `devtools::load_all()` | allow you to use package & see how things are going | 


- ensure has R package development correctly setup
```{r}
devtools::has_devel()
```

## Build my first R package
```{r eval=FALSE}
devtools::create("~/tmp/yoh")

# Creating package yoh in /Users/csiu/tmp
# No DESCRIPTION found. Creating with values:
# 
# 
# Package: yoh
# Title: What the Package Does (one line, title case)
# Version: 0.0.0.9000
# Authors@R: person("First", "Last", email = "first.last@example.com", role = c("aut", "cre"))
# Description: What the package does (one paragraph).
# Depends: R (>= 3.2.2)
# License: What license is it under?
# LazyData: true
# Adding RStudio project file to yoh
```

File layout
```
[Tue Nov 10 10:46:22]~/tmp/yoh/$ tree
-- .
   |-- .gitignore
   |-- .Rbuildignore
   |-- DESCRIPTION
   |-- NAMESPACE
   |-- R
   |-- yoh.Rproj
```

@DESCRIPTION
```
Package: yoh
Title: What the Package Does (one line, title case)
Version: 0.0.0.9000
Authors@R: person("First", "Last", email = "first.last@example.com", role = c("aut", "cre"))
Description: What the package does (one paragraph).
Depends: R (>= 3.2.2)
License: What license is it under?
LazyData: true
```

@NAMESPACE
```
exportPattern("^[^\\.]")
```

- first thing to do (after opening RProject in RStudio) put under version control (then need to restart RStudio)
```{r eval=FALSE}
devtools::use_git()
```
