---
title: "tidy data & reshape"
author: "csiu"
date: "October 15, 2015"
output: 
    html_document:
        keep_md: yes
---

- [lecture outline](http://stat545-ubc.github.io/cm012_tidy-data-factors.html)

```{r}
suppressPackageStartupMessages(library(dplyr))
```

- [examples of good peer reviews and homework](https://github.com/STAT545-UBC/2015_pvt-to-students-and-staff)

- [Be the boss of your factors](http://stat545-ubc.github.io/block014_factors.html)
### Pass arguments as strings
- `aes_string()` ... allows you to pass string in
- `summarise(mtcars, mean(mpg))` is the same as `summarise_(mtcars, "mean(mpg)")` -- not the underscore

--- 

## Tidying data
### The dataset
- `EDAWR` - r packages with all of the data sets that will be used today
```{r}
# install.packages("devtools")
# devtools::install_github("rstudio/EDAWR")

library(EDAWR)
?storms
?cases
?pollution
```

### The package
- `tidyr` (package for tidying data) is a subset of `reshape2`

```{r}
library(tidyr)
```

- `tidyr::gather()` - need to specify KEY and VALUE variable
- `tidyr::spread()` 
    - specify which is the KEY variable ... destined to be new column names 
    - specify which is the VALUE variable ... destined to be values

### Example of `gather()`
```{r}
cases
cases %>% gather(year, n, -country)
gather(cases, "year", "n", 2:4) ##2:4 <- which columns you want to stack up
gather(cases, "year", "n", one_of("2011", "2012", "2013"))
```

### Example of `spread()`
```{r}
pollution

spread(pollution, "size", "amount")
```

### Other handy functions: `separate()` and `unite()`
- `tidyr::separate()`
```{r}
storms
(storms2 <- separate(storms, date, c("year", "month", "day"), sep="-"))
```

- `tidyr::unite()`
```{r}
unite(storms2, date, year, month, day, sep="-")
```

----
## Factor level order 
- [factor management](http://stat545-ubc.github.io/block014_factors.html)
    - `droplevels()` to drop unused factor levels
    - `reorder()` to reorder factor levels
    - `revalue()`
    - growing a factor
    - creating a factor from scratch: `factor()`
